---
title: "fire_STSM"
author: ""
date: "12 December 2017"
output: pdf_document
---

# Fire state-and-transition toy model

This module contains a VERY simple state-and-transition fire model.
For now, it is a stand-alone module, that uses the LCC 2005 vegetation map and relies on the `SpaDES::spread2` to calculate fire spread.

# Usage

```{r module_usage}
library(igraph)
library(raster)
library(SpaDES)

moduleDir <- file.path("C:/Ceres/GitHub/LandscapesInMotion/R/SpaDES/m")
inputDir <- file.path("C:/Ceres/GitHub/LandscapesInMotion/R/SpaDES/inputs")
outputDir <- file.path("C:/Ceres/GitHub/LandscapesInMotion/R/SpaDES/outputs")
cacheDir <- file.path("C:/Ceres/GitHub/LandscapesInMotion/R/SpaDES/cache")

## Get foothills study area
foothills <- shapefile("C:/Ceres/GitHub/LandscapesInMotion/R/SpaDES/inputs/Alberta_study_area/Alberta_study_area")
LCC2005 <- raster("C:/Ceres/GitHub/LandscapesInMotion/R/SpaDES/inputs/LCC2005_V1_4a.tif")

## simulation parameters
paths <- list(
  cachePath = cacheDir,
  modulePath = moduleDir,
  inputPath = inputDir,
  outputPath = outputDir
)

modules <- list("fire_STSM")

times <- list(start = 0, end = 10)

parameters <- list(
  fire_STSM = list(.useCache = TRUE, fireSize = 1000, noStartPix = 100)
)

objects <- list("studyArea" = foothills, "LCC2005" = LCC2005)

mySim <- simInit(times = times, params = parameters, modules = modules,
                 objects = objects, paths = paths)

dev()
clearPlot()
spades(mySim, debug = TRUE)
```

# Events

## Initialisation

The module is initialised by cropping and reprojecting the full LCC 2005 vegetation raster to a smaller study area (which is now provided externally in the global, or Rmd scripts). 

The LCC 2005 classes are then converted to 5 broader classes: 
* **Non-burnable**. Vegetation classes that do not burn. Comprises: wetlands (19), wet tundra (23) cropland/woodlands (26, 27,28,29), lichen dominated habitats (30, 31,32), rock outcrops (33), recent burns (34), cities (36), water (37, 38), snow/ice (39)
* **Grasslands**. Open habitats that burn easily but at low intensities. Comprises classes 17, 18, 20, 21, 22, 24, 25
* **Shrublands**. Shrub dominated habitats that burn easily at medium intensity. Comprises: old burns (i.e. forest that burned in the last 10y; 35), shrublands (16)
* **Deciduous forest**. Forests dominated by deciduous trees. Comprises classes 2, 11, 12
* **Mixed forest**. Forests of deciduos and conifer species. Comprises classes 3, 4, 5, 13, 14, 15
* **Coniferous forest**. Conifer-dominated forests. Comprises classes 1, 6, 7, 8, 9, 10.

The conversion of LCC classes results in a `vegetation_prefire` raster.

During the initialisation, raster maps of `burnable_areas` and rate-of-spread (`ROS_map`) are created based on the `vegetation_prefire` raster:
* `burnable_areas` is identical to `vegetation_prefire` except in that NAs are attributed to all "non-burnable" pixels.
* `ROS_map` contains the relative spread probabilities (probability of "catching fire" from a neighbour pixel) for each class. These are:

```{r ROS_table echo = FALSE}
knitr::kable(data.frame(Class = c("non-burnable", "grasslands", "shrublands", "deciduous forest", "mixed forest", "coniferous forest"),
                        ROS = c(0, 0.9, 0.7, 0.2, 0.4, 0.5)), caption = "Rates of spread for vegetation classes")
```

## Fire spread

Fire spread is calculated using the `SpaDES::spread2` function. This function makes use of the `burnable_areas` and `ROS_map` rasters to calculated where fire spreads to after it is "seeded" in the start pixels (`startPix`). The number of starting pixels is defined as a module parameter (`noStartPix`) and drawn at random from `burnable_areas` raster. Fire sizes (no. of pixels ?) are also determined externally using the parameter `fireSize`.

## Vegetation transitions and calculation of fire severity

After fire map is generated (`spreadRas`) vegetation state transitions are calculated for pixels that burned (that have fire) based on a transition probability matrix (`fire_transitprobs`).
* Grasslands and shrbulands stay/become grasslands
* Deciduous and mixed forests become shrublands
* Conifer forests become mixed forests

Fire severity is then  calculated based on the type of vegetation transitions that occurred after fire:
* **Low severity** when grasslands burn
* **Moderate severity** when shrublands burn
* **High severity** when any type of forest burns

